Our Future Nuclear Data Needs

### 2018-12-16 nuclear_data_needs/nuclear_data_needs.tex
Merged in Lee's comments.  Added conclusion section and the title page keywords.

%Keywords, etc.
\begin{keywords}
\textred{Add keywords here!}  keywords, separated by comma, no full stop, lowercase
\end{keywords}
\maketitle

@@

\begin{wrapfigure}{r}{0.4\textwidth}
\centering
\includegraphics[width=0.39\textwidth]{pipeline_cropped.pdf}
\caption{\label{fig:pipeline}The nuclear data pipeline.}
\end{wrapfigure}

@@

The ideal way to improve (n,n') data on the \ce{^{235,238}U} and \ce{^{239}Pu} would be to perform measurements of the scattered neutron in coincidence with transitions emitted by the excited nucleus.  In the case of the Actinides this will potentially require the observation of conversion electrons given that the lowest transitions in all actinides are almost entirely converted.  Unfortunately, neutron beams usually have quite low intensity, with a fluence in the fast (1--10 MeV) region between $10^3$ and $10^7$ n/s/cm$^2$ at the target location.  Furthermore, many spallation neutron sources have a significant high-energy spectral component that can cause inter-element scattering in a neutron scintillator array rendering precise determination of the scattered neutron energy difficult to  perform. However, intense limited-energy neutron sources based on thick-target deuteron breakup \cite{Harrig2018} and forward-fitting techniques developed for use at spallation neutron sources \cite{Kee18} off an opportunity to perform neutron-gamma coincident measurements in the fast fission energy spectrum.  These techniques could provide valuable data to improve nuclear data evaluations of inelastic scattering, even on targets where fission is an open channel.

@@

The NDIWG process reflects a growing recognition that improved nuclear data is essential to a wide range of human endeavors that transcend the needs of any one societal need or governmental mission.  Its goal is the development of a new, multi-agency national plan to address nuclear data needs in a coherent way.  This plan, which is being developed in partnership with international collaborators at the IAEA and the OECD-NEA represents a new paradigm for collaborative research.  
% 
% %Heading 1
% \section{FIRST-LEVEL HEADING}
% This is dummy text. 
% % Heading 2
% \subsection{Second-Level Heading}
% This is dummy text. This is dummy text. This is dummy text. This is dummy text.
% 
% % Heading 3
% \subsubsection{Third-Level Heading}
% This is dummy text. This is dummy text. This is dummy text. This is dummy text. 
% 
% % Heading 4
% \paragraph{Fourth-Level Heading} Fourth-level headings are placed as part of the paragraph.

>>>>

%Keywords, etc.
\begin{keywords}
nuclear data,  structure,  reactions, evaluation, fission, medical Isotopes, nuclear energy, nonproliferation, inelastic scattering, level density, radiative strength, $\beta$-decay, reactor antineutrinos
\end{keywords}
\maketitle

@@

\begin{wrapfigure}{r}{0.39\textwidth}
\centering
\includegraphics[width=0.38\textwidth]{pipeline_cropped.pdf}
\caption{\label{fig:pipeline}The nuclear data pipeline.}
\end{wrapfigure}

@@

The ideal way to improve (n,n') data on the \ce{^{235,238}U} and \ce{^{239}Pu} would be to perform measurements of the scattered neutron in coincidence with transitions emitted by the excited nucleus.  In the case of the Actinides this will potentially require the observation of conversion electrons given that the lowest transitions in all actinides are almost entirely converted.  Unfortunately, neutron beams usually have quite low intensity, with a fluence in the fast (1--10 MeV) region between $10^3$ and $10^7$ n/s/cm$^2$ at the target location.  Furthermore, many spallation neutron sources have a significant high-energy spectral component that can cause inter-element scattering in a neutron scintillator array rendering precise determination of the scattered neutron energy difficult to  perform. However, intense limited-energy neutron sources based on thick-target deuteron breakup \cite{Harrig2018} and forward-fitting techniques developed for use at spallation neutron sources \cite{Kee18} offer an opportunity to perform neutron-gamma coincident measurements in the fast fission energy spectrum.  These techniques could provide valuable data to improve nuclear data evaluations of inelastic scattering, even on targets where fission is an open channel.

@@

The NDIAWG process reflects a growing recognition that improved nuclear data is essential to a wide range of human endeavors that transcend the needs of any one societal need or governmental mission.  Its goal is the development of a new, multi-agency national plan to address nuclear data needs in a coherent way.  This plan, which is being developed in partnership with international collaborators at the IAEA and the OECD-NEA represents a new paradigm for collaborative research.

\section{CONCLUSION \texorpdfstring{---}{-} RETAINING AND TRAINING A SKILLED NUCLEAR DATA WORKFORCE}

A robust nuclear data evaluation process is needed to provide the information required for many applications ranging from nonproliferation to nuclear energy to isotope production.  However, it takes many years of post-Ph.D. experience to train nuclear data evaluators, and only a short break in support to lose  this work force to other pursuits.  It is therefore critical for the many societal programs that rely on nuclear data to support individuals engaged in maintaining all functions that make up the \enquote{Nuclear Data Pipeline}, from measuring and modeling through compilation, evaluation, validation, and processing.  As more and more of the research in the United States moves over to a project-driven model, it is imperative that sponsors coordinate their efforts to ensure that there is no loss of talent in the field.  The NDIAWG discussed above is a significant step to help address these issues in a systematic fashion.

Furthermore, nuclear data evaluation is an \emph{aging} field, with the mean age of an evaluator in the US being well-over 50.  It is essential that the next generation of nuclear data evaluators be recruited and trained while the current set of experienced individuals are still active in the field.  This is particularly challenging since there is, \emph{per se}, no graduate education program in nuclear data evaluation, and the vast majority of the centers with evaluation expertise are not associated with degree-granting institutions.

Perhaps one of the \enquote{endangered species} in nuclear data is the experimental nuclear data evaluator.  This isn't an experimentalist doing a new, better experiment, nor a data library creator that makes an ENDF-like library, but a person who knows which of the more than 22,000 measurements in the EXFOR/CSISRS library are to be taken seriously, discarded or normalized to newer standards.  This sort of a \enquote{hybrid} individual would be capable of providing not only good mean values, but also the well-quantified uncertainties (including covariances) that are needed for most applications.  A concerted effort on the part of both academia and the international research community is needed to ensure that talented early-career individuals with the right combination of interest and inclination be afforded the opportunity to get trained by an \enquote{international village} of evaluators and application subject matter experts.

In closing, it is worth noting that the dramatic increase in computational capabilities over the past several decades allows for the sort of detailed calculations needed to fulfill the promise of many applications, \emph{provided that the correct evaluated data is available}.   Recent efforts in the US offer the promise of a national nuclear data plan to provide the needed data.  If these efforts are maintained it could help herald in a world with a new sense of safety from enhanced nonproliferation capabilities, a new source of safe, carbon-neutral energy and a new class of radio-pharmaceuticals to diagnose and treat illness.

% 
% %Heading 1
% \section{FIRST-LEVEL HEADING}
% This is dummy text. 
% % Heading 2
% \subsection{Second-Level Heading}
% This is dummy text. This is dummy text. This is dummy text. This is dummy text.
% 
% % Heading 3
% \subsubsection{Third-Level Heading}
% This is dummy text. This is dummy text. This is dummy text. This is dummy text. 
% 
% % Heading 4
% \paragraph{Fourth-Level Heading} Fourth-level headings are placed as part of the paragraph.
